---
title: "Energy Price and Demand app"
author: "Group cdjs (Chris, David, Jaehee, Santiago)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Use Case

This app has been designed in order to explore the differences in energy systems across different states and over time. This app may be used by businesses and economists who may want to analyze both the "outputs" of the energy system, like the total amount of electricity produced or the total carbon emissions generated, as well as the "inputs" of the energy system, like the average price of electricity or the number of customer accounts. In addition, we know that for some kinds of businesses, like the manufacturing sector, a large fraction of operating costs pertain to energy alone; for these individuals, it would be useful to be able to compare electricity prices both over time and over geography to understand 1) what their future costs might look like in the states they are currently located in, and 2) where might be the best places to build other factories based on electricity costs.

### Who is the expected typical user? 

As mentioned above, this was primarily designed for those with a stake in the energy system. This might be someone on the regulatory side, such as an employee of the Federal Energy Regulatory Commission or a state lawmaker, or it could also be someone on the business side, like a factory owner interested in understanding their future costs. However, we also built in a few features that could be interesting for students or everyday Americans interested in the energy sector. Since we don't "see" -- or even think about -- the energy system very often, this audience might be interested in learning that electricity prices vary across the day, and might be interested in visualizing that relationship.

### Typical questions that this app should help the user answer.

* Question 1 - Which states have the highest and lowest electricity costs? Have they always, or is this something new?

* Question 2 - How does electricity demand vary depending on the time of day, and do different states have different patterns of demand?

* Question 3 - How do a state's carbon emissions relate to its price of electricity? How has that changed over time?

# Required Packages

## What Packages to Download

The use of the app requires that the user must have the following packages installed:

## Description of the required packages
* `eia`- eia provides API access to data from the US Energy Information Administraion
* `shiny` - shiny helps to build interactive web applications with R
* `tidyverse` -  tidyverse is an opinionated collection of R packages
* `lubridate`- lubridate helps to make it easier to manipulate and parse dates.
* `ggcorrplot` - ggcorrplot is a graphical display of a correlation matrix using ggplot2
* `broom`- broom can convert statistical analysis objects from R into tidy tibbles
* `shinythemes` - shinythemes helps to change the overall appearance of the Shiny application

## Version used in the app
* `eia` - In order to run the app the use of version >=0.3.7 is suggested and >=0.1.0 is required.  
* `shiny` - In order to run the app the use of version >= 1.0.0 is required.  
* `tidyverse` - In order to run the app the use of version >= 3.3.3 is required.  
* `lubridate` - In order to run the app the use of version >= 1.7.1 is suggested and >=1.7.0 is required.  
* `ggcorrplot` - In order to run the app the use of version >= 0.1.3 is suggested.  
* `broom` - In order to run the app the use of version >=0.3 is required. 
* `shinythemes` - In order to run the app the use of version >=1.1.1.  


# Data Source and Structure

 
## Data Source 
* Data was collected from Energy Information Administration(EIA) -> https://www.eia.gov/opendata/ 

* Get access to EIA API query browser ->  https://www.eia.gov/opendata/qb.php

* For the main data, we track five variables by year:

    * Average retail electricity price
    * Number of customer accounts 
    * Carbon emissions
    * Net electricity generation
    * Revenue from retail sales of electricity
    
* We also track the demand for electricity in each grid region, although this is by hour.

* Here is the guideline to find each data from the API query browser. After navigating to the page, the user will be able to see the series prefix used in the API calls. 

    * Average retail electricity price: EIA Data Sets -> Electricity -> Average retail price of electricity -> Industrial
    * Number of customer accounts: EIA Data Sets -> Electricity -> Number of customer accounts
    * Carbon Emissions: EIA Data Sets -> CO2 Emissions -> by State
    * Net Electricity generation: EIA Data Sets -> Electricity -> Net generation
    * Revenue from retail sales of electricity: EIA Data Sets -> Electricity -> Retail sales of electricity

* We use the package "eia" as a wrapper to facilitate our api calls to the EIA database.

* There were yearly, monthly, quarterly for some, but not all, data. As a result, this app interpolates data to track by year, the most granular timeframe all variables have in common. The exception is daily demand, which is tracked separately by hour.

## Data Structure
After cleaning the data, we combined them into one dataframe which is Full_data
```{r}
library(readr)
Full_data <- readRDS("./data/yearly_data.rds")
str(Full_data)
```


# EDA Inputs, Controls, and Outputs

In order to perform EDA, the data needed to be cleaned. Here is the explanation on how each tab get along with the app and will allow the user to conduct an effective EDA.

## Tab 1: Univariate Analysis
The **univariate** tab allows the user to see the distribution for each of the variable. This allows the user to see how each variable is distributed by state or year. 

1. Filter by State or Year: 
    * User can interactively choose between state or year. Because the data is time-series, we have to filter once (as we do here) before we can plot distributions.

1. Select State or Year: 
    * If the user chose state, then this filter will automatically give the users options on 50 states. 
    * If the user chose year, then this filter will automatically give the users options on year (2008-2020)

1. Choose Variable: 
    * The variable of interest

1. Choose a plot type:
    * This tab allows the user to choose between a "Density Plot", "Histogram", "Frequency Polygon", in order to make the user may choose the type of the plot to visualize better/preferred.

1. Number of Bins for Histogram: 
    * Users can interactively change the number of bins (min =1, max = 100) in the histogram.
    
1. T-test Null Value:
    * Users can select the null value against which their selected variable will be compared in a t-test.

1. t.test? 
    * Output the results of the one-sample t.test. The results of the test changes depending on the variables(electricity_price, carbon_emissions, customers, retail_sales, total_electricity) that users choose.

1. NB
    * Input at the end to give additional explanation on the data. After cleaning the data, we found out that there is no data for 'carbon emissions'(2018 ~ 2020) & 'customers'(2001 ~ 2007). So if the users choose these options they cannot see the graph. 

## Tab 2: Bivariate Analysis
The **bivariate** tab allows the user to make comparisons between two variables in the data set. 
[PLOT 1] will help users understand the relationship between two variables by one of the states/by one of the years that users chose
[PLOT 2] will help users understand the relationship between two variables by based on all the data

1. Filter state or year? 
    * User can interactively choose between state or year. 

1. Filter which state? Filter what year? 
    * If the user chose state, then this filter will automatically give the users options on 50 states. 
    * If the user chose year, then this filter will automatically give the users options on year(2001-2020)

1. Variables?
    * The variable of interest for the X axis and the Y axis 

1. Whether to add an OLS line in [PLOT 1]? 
    * Allows the user to see a display of linear regression model of [PLOT 1], so that the user may make analyze based of the data. 

1. All states? All year? 
    * This checkboxInput will help user to see all the data regarding two variables(X axis, Y axis) that the user chose. Also, it will be colored by state. 

1. Wheter to add an OLS line in [PLOT 2]? 
    * Allows the user to see a display of linear regression model of [PLOT 2], so that the user may make analyze based of the data.

1. Show summary? 
    * If the 'Show summary?' is requested, add a summary below [PLOT 2] from running a linear model based on [PLOT2].

1. Warning
    * Input at the end to give additional explanation on the data. After cleaning the data, we found out that there is no data for 'carbon emissions'(2018 ~ 2020) & 'customers'(2001 ~ 2007). So if the users choose these options they cannot see the graph.


## Tab 3: Multivariate Analysis 
The **multivariate** tab allows the user to see analysis based on more than two variables.This tab will provide the users better understanding of the relationships between the several variables. Then, users can see in one glance all of the relationships and compare which one has strong positive correlation or negative correlation. 

Since there is no data for 'carbon emissions'(2018 ~ 2020) & 'customers'(2001 ~ 2007), the rows(observations) of all five variables doesn't match. Therefore, we selected the overlapping years from 2008~2017 in order to make the rows same to conduct correlation and pairs plot.   

1. Correlation 
    * This tab will output the correlation matrix heatmap from 2008~ 2017 across all states. 
If the number is close to 1, it will be strong positive correlation, but if the number is close to -1, it will be strong negative correlation. This is not interactive plot with the users but will provide overall correlations between the variables. 

1. Pairs
    * This tab will output the scatter plot matrices from 2008 ~ 2017 across all states.
Shows all of the scatter plots between variables in one glance, also this is not interactive plot, but will provide overall relationships between the variables. 


## Tab 4: Daily Load
This tab will help users to compare two regions based on demand for electricity throughout the day. This tab used local time variable rather than UTC

1. Choose 1? 
    * Select one region from California, Carolina, Central, Florida, Mid-Atlantic, Midwest, New England, New York, Northwest, Southeast, Southwest,Tennessee, Texas

1. Choose 2?
    * Select another region from California, Carolina, Central, Florida, Mid-Atlantic, Midwest, New England, New York, Northwest, Southeast, Southwest,Tennessee, Texas

1. Date to Compare?
    * This will help users to choose the date that they want to see.

## Tab 5: Time Series
This tab allows user to interactively see the trend of variables. Four plots in this tab X-axis is automatically year variable(time), since time series plot is dealing with the trend over time. 
User can choose a state first, and then choose two variables to compare trends. PLOT 3 and PLOT 4 will show the users the trend of the variables over time regarding all of the states data.

1. Choose state? 
    * User can choose from 50 states

1. What variable trend do you want to see(choice 1)?
    * User can choose from 'Average retail electricity price', 'carbon emissions', 'number of customer accounts', 'revenue from retail sales of electricity', 'net electricity generation'

1. What variable trend do you want to see(choice 2)? 
    * User can choose from 'Average retail electricity price', 'carbon emissions', 'number of customer accounts', 'revenue from retail sales of electricity', 'net electricity generation'

1. Add trend smooth line?
    * This request to show trend of the variable that user chose

1. All states? 
    * Show the trend of the variable regarding all of the data regardless of state filter. 
    * User can compare two variables trend with PLOT 3 and PLOT 4

## Tab 6: Spreadsheet

This tab will show all of the data
Users can interact with this spreadsheet. They can search or choose entries that they want to see. 


# How to use interactively between tabs?

For example, the user may see that there is a trend that carbon emissions are going up with time in Time series tab. Then the user may switch to the Bivariate tab in order to see how energy prices in that state correlate with an increase in carbon emissions. 

# References

There are three references from either academic papers or widely available authoritative sources that are relevant to this app. See below: 


1. Dhawan, R., & Jeske, K. (2006). How resilient is the modern economy to energy price shocks? Economic Review (Atlanta, Ga.), 91(3), 21–.

1. Edelstein, P., &amp; Kilian, L. (2007, November 7). The Response of Business Fixed Investment to Changes in Energy Prices: A Test of Some Hypotheses about the Transmission of Energy Price Shocks. De Gruyter. https://www.degruyter.com/document/doi/10.2202/1935-1690.1607/html. 

1. Kilian, L. (n.d.). The Economic Effects of Energy Price Shocks. Journal of Economic Literature. https://www.aeaweb.org/articles?id=10.1257%2Fjel.46.4.871. 











