---
title: "Energy Price and Demand app"
author: "Group cdjs (Chris,David,Jaehee,Santiago)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Use Case


This app has been designed in order to analyze the demand as well as the price for energy inputs across different states. This app may be used by businesses and economists who may want to analyze the usage of different sources of energy as well as the price by different states. We know that for the manufacturing sector as well as other business sectors, a large fraction of costs pertain to energy alone and hence, this app uses data from the eia in order to provide those who want to make energy related business decisions with statistical and graphical information.   


## Who is the expected typical user? 

This app may be used by business analysts, economists, policymakers, or general public who may want to analyze the usage of different sources of energy as well as the price by different states. 
We know that for the manufacturing sector as well as the business sector, a large fraction of costs pertain to energy alone. Therefore, this app uses data from the EIA in order to provide those who want to make energy related business decisions, policymaking, or public understanding. This app will help improve the understanding of the interaction with the economy and environment.

## Typical questions that this app should help the user answer.

Question 1 - 

Question 2 - 

Question 3 - 

# Required Packages

## What Packages to Download

The use of the app requires that the user must have the following packages installed:

## Description of the required packages
1. `eia`- eia provides API access to data from the US Energy Information Administraion
2. `shiny` - shiny helps to build interactive web applications with R
3. `tidyverse` -  tidyverse is an opinionated collection of R packages
4. `lubridate`- lubridate helps to make it easier to manipulate and parse dates.
5. `ggplot2` - ggplot2 helps to create graphics
6. `ggthemes` - ggthemes provides extra themes such as theme_bw()
7. `ggcorrplot` - ggcorrplot is a graphical display of a correlation matrix using ggplot2
8. `GGally`- GGally is an extension to ggplot2 and it includes pairwise plot matrix
9. `broom`- broom can convert statistical analysis objects from R into tidy tibbles
10. `shinythemes` - shinythemes helps to change the overall appearance of the Shiny application

## Version used in the app
1. `eia` - In order to run the app the use of version >=0.3.7 is suggested and >=0.1.0 is required.  
2. `shiny` - In order to run the app the use of version >= 1.0.0 is required.  
3. `tidyverse` - In order to run the app the use of version >= 3.3.3 is required.  
4. `lubridate` - In order to run the app the use of version >= 1.7.1 is suggested and >=1.7.0 is required.  
5. `ggplot2` - In order to run the app the use of version >= 3.0.0 is required.  
6. `ggthemes` - In order to run the app the use of version >= .  
7. `ggcorrplot` - In order to run the app the use of version >= 0.1.3 is suggested.  
8. `GGally` - In order to run the app the use of version >= . 
9. `broom` - In order to run the app the use of version >=0.3 is required. 
10. `shinythemes` - In order to run the app the use of version >=1.1.1.  


# Data Source and Structure

 

## Data Source 
Data was collected from Energy Information Administration(EIA) -> https://www.eia.gov/opendata/ 

Get access to EIA API query browser ->  https://www.eia.gov/opendata/qb.php

Five Data that is used in this app are: 

Electricity Price, Customers, Carbon Emissions, Total Electricity, Retail Sales.

In order to find each data fast, here is the guideline to find each data from the API query browser: 

 1. Electricity Price: EIA Data Sets -> Electricity -> Average retail price of electricity -> Industrial
 2. Customers: EIA Data Sets -> Electricity -> Number of customer accounts
 3. Carbon Emissions: EIA Data Sets -> CO2 Emissions -> by State
 4. Total Electricity: EIA Data Sets -> Electricity -> Net generation
 5. Retail Sales: EIA Data Sets -> Electricity -> Retail sales of electricity

Use eia package to have an API access to data from EIA 

There were yearly, monthly, quarterly for each data but this app only uses yearly data

## Data Structure
After cleaning the data, we combined them into one dataframe which is Full_data
```{r}
library(readr)
Full_data <- read_csv("Full_data.csv")
str(Full_data)
```



# EDA Inputs, Controls, and Outputs

In order to perform EDA, the data needed to be cleaned. Here is the explanation on how each tab get along with the app and will allow the user to conduct an effective EDA.

## Tab 1: Univariate Analysis
The **univariate** tab allows the user to see the distribution for each of the variable. This allows the user to see how different energy sources or energy prices are distributed in each state/year. 

### Filter state or year? 
User can interactively choose between state or year. 

### Filter which state? Filter what year? 
If the user chose state, then this filter will automatically give the users options on 50 states. 
If the user chose year, then this filter will automatically give the users options on year(2001-2020)

### Variable? 
The variable of interest

### Choose a plot type?
This tab allows the user to choose between a "Density Plot", "Histogram", "Frequency Polygon", in order to make the user may choose the type of the plot to visualize better/preferred

### Histogram: Number of Bins? 
Users can interactively change the number of bins(min =1, max = 100) in the histogram if they want 

### t.test? 
Output the results of the one-sample t.test. The results of the test changes depending on the variables(electricity_price, carbon_emissions, customers, retail_sales, total_electricity) that users choose.

### Warning
Input at the end to give additional explanation on the data. After cleaning the data, we found out that there is no data for 'carbon emissions'(2018 ~ 2020) & 'customers'(2001 ~ 2007). So if the users choose these options they cannot see the graph. 


## Tab 2: Bivariate Analysis
The **bivariate** tab allows the user to make comparisons between two variables in the data set. 

### Filter state or year? 
User can interactively choose between state or year. 

### Filter which state? Filter what year? 
If the user chose state, then this filter will automatically give the users options on 50 states. 
If the user chose year, then this filter will automatically give the users options on year(2001-2020)

### Variables?
The variable of interest for the X axis and the Y axis 

### Whether to add an OLS line in [PLOT 1]? 
Allows the user to see a display of linear regression model of [PLOT 1], so that the user may make analyze based of the data. 

### All states? All year? 
This checkboxInput will help user to see all the data regarding two variables(X axis, Y axis) that the user chose. Also, it will be colored by state. 

### Wheter to add an OLS line in [PLOT 2]? 
Allows the user to see a display of linear regression model of [PLOT 2], so that the user may make analyze based of the data.

## Show summary? 
If the 'Show summary?' is requested, add a summary below [PLOT 2] from running a linear model based on [PLOT2].

## Tab 3: Multivariate Analysis 
The **multivariate** tab allows the user 


## Tab 4: Daily Load


## Tab 5: Time Series

## Tab 6: Spreadsheet

    Data cleaning
    Initial / descriptive data analyses
    Question finding / hypothesis generation

### Example of how to use?
For example, let's say that the user chose for filters state, California and chose X axis for carbon emissions and Y axis for electricity price

For example, the user may see that there is trend that carbon emissions are going up with time. The user may then switch to the bivariate tab in order to see how energy prices in that state correlate with an increase in carbon emissions. 

#3 output

# Refrences

There are three references from either academic papers or widely available authoritative sources that are relevant to this app. See below: 


1. https://www.frbatlanta.org/-/media/documents/research/publications/economic-review/2006/vol91no3_dhawan-jeske.pdf

2. https://www.degruyter.com/document/doi/10.2202/1935-1690.1607/html

3. https://www.aeaweb.org/articles?id=10.1257/jel.46.4.871






# >>>>>>> 677de5201203db6a00bd8bbf7737318e60c7c7fc








